### 2014-04-22 docs/paper/paper.tex
real vs shared PCs

Power analysis showed that the implemented design, with 7 one-dimensional and 7 two-dimensional population units,
is estimated to use a maximum of 5.675~W of power while a simulation is running, assuming all population units are fully utilized.
With the installed heatsink and fan on the VC707 board, the FPGA is expected to operate at an on-die temperature of $31.5\,^{\circ}{\rm C}$.

@@

A backend to Nengo performs the task of compiling a Nengo model to a loadfile that can be
simulated on the hardware. The compiler calculates principal components for each population
of neurons, and then performs a clustering operation in order to associate many
populations with each set of principal components while minimizing the differences
between the approximate principal components and the actual values. % FIXME explain/calculate why actual principal components cannot be used
The compiler calculates decoders and encoders for each population with respect
to the shared principal components, and computes an optimized encoder schedule % FIXME details needed here
that coordinates reads made by each encoder so that all operations can be performed
in as few clock cycles as possible while avoiding a read collision on any decoded value block.
Once the compiler has finished writing out the loadfile, the Nengo simulator
transfers the configuration data to the hardware and controls input and output during the simulation.

@@

\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot
% FIXME figure: larger axis and label font, make this consistent across all figures

@@

The oscillator requires a brief non-zero ``start-up'' input in order to begin oscillating.
After this input disappears (0.1~s), both dimensions of the decoded output can be seen to
oscillate in a sinusoidal fashion. This oscillation, once started, will proceed indefinitely.
% FIXME frequency analysis?
The oscillation as observed in the figure confirms the correct operation of the two-dimensional
population hardware.

@@

% FIXME: explain input and expected output

@@

At present, the interconnect between the encoders and the decoded value
buffers is the most complex part of the design in terms of the density of its logic
in hardware. If there are $N$ encoders and $M$ decoded value buffers, a total of
$NM$ connections must be made between these components in order for every
encoder to be able to read from every decoded value buffer.
Note that this is still a massive improvement over a neuron-to-neuron interconnect,
which would require over 10000 times as many connections to provide all-to-all connectivity.
% FIXME contrast this with size of neuron-to-neuron interconnect
For large numbers
of encoders and decoded value buffers, the complexity of the interconnect at
that design size makes it impossible for the FPGA design tools to find a way to
implement that part of the design on the device. This is currently the limiting
factor in terms of the number of population units that can be implemented on
the development board (as opposed to, for example, a lack of memory blocks for
decoded value buffers). A future version of the \design{} design may address
this issue and implement an interconnect that scales well to larger numbers of
encoders and decoded value buffers in order to allow more populations to be
simulated. % FIXME how could this be done?

>>>>

Power analysis showed that the implemented design, with 7 one-dimensional and 7 two-dimensional population units,
is estimated to use a maximum of 5.675~W of power while a simulation is running, assuming all population units are fully utilized.
For an individual population unit, a one-dimensional unit consumes an average of 0.130~W of power during simulation,
and a two-dimensional unit consumes an average of 0.425~W of power.
With the installed heatsink and fan on the VC707 board, the FPGA is expected to operate at an on-die temperature of $31.5\,^{\circ}{\rm C}$.

@@

A backend to Nengo performs the task of compiling a Nengo model to a loadfile that can be
simulated on the hardware. The compiler calculates principal components for each population
of neurons, and then performs a clustering operation in order to associate many
populations with each set of principal components while minimizing the differences
between the approximate principal components and the actual values.

Populations must share principal components because storing a separate set of principal components for each population would require
an unreasonably large amount of memory. As an example, a one-dimensional population currently uses seven principal components,
each of which is a 1024-word, 12-bit block memory, for a total of 86~016~bits of memory per population.
The FPGA on the VC707 development board has a maximum of approximately 37~080~000~bits of block memory;
if we allowed each population to use its own principal components, we could not simulate more than 431 populations this way
(and this number is halved for two-dimensional populations, which use 15 principal components).
Therefore, in order to simulate large numbers of populations, it is necessary for large groups of populations to share principal components,
even though this is not as accurate as providing a ``real'' set of principal components for each separate population.

The compiler calculates decoders and encoders for each population with respect
to the shared principal components, and computes an optimized encoder schedule % FIXME details needed here
that coordinates reads made by each encoder so that all operations can be performed
in as few clock cycles as possible while avoiding a read collision on any decoded value block.
Once the compiler has finished writing out the loadfile, the Nengo simulator
transfers the configuration data to the hardware and controls input and output during the simulation.

@@

\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot (for all figures)
% FIXME figure: larger axis and label font, make this consistent across all figures

@@

The oscillator requires a brief non-zero ``start-up'' input in order to begin oscillating.
After this input disappears (0.1~s), both dimensions of the decoded output can be seen to
oscillate in a sinusoidal fashion. This oscillation, once started, will proceed indefinitely.
The oscillation as observed in the figure confirms the correct operation of the two-dimensional
population hardware.

@@

@@

At present, the interconnect between the encoders and the decoded value
buffers is the most complex part of the design in terms of the density of its logic
in hardware. If there are $N$ encoders and $M$ decoded value buffers, a total of
$NM$ connections must be made between these components in order for every
encoder to be able to read from every decoded value buffer.
Note that this is still a massive improvement over a neuron-to-neuron interconnect,
which would require over 10000 times as many connections to provide all-to-all connectivity.
For large numbers of encoders and decoded value buffers, the complexity of the interconnect at
that design size makes it impossible for the FPGA design tools to find a way to
implement that part of the design on the device. This is currently the limiting
factor in terms of the number of population units that can be implemented on
the development board (as opposed to, for example, a lack of memory blocks for
decoded value buffers). A future version of the \design{} design may address
this issue and implement an interconnect that scales well to larger numbers of
encoders and decoded value buffers in order to allow more populations to be
simulated. % FIXME how could this be done?
